% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.

\chapter{Introduction}\label{chapter:introduction}

Studying seismic phenomena provides valuable insights and potential techniques for identifying areas that are more susceptible to earthquakes, mitigating economic impacts, 
and reducing casualities. \\

The field of earthquake simulation has seen significant advancements in recent years, thanks to the availability of powerful supercomputers ~\parencite{inproceedings}, ~\parencite{breuer}, ~\parencite{seissol}.
With these resources, it has become possible to model realistic 3D earthquake events on a large scale, something that was not feasible until a few decades ago. 
Today's simulation software heavily relies on numerical methods to approximate real-life physics. In this process, the volume of the problem is discretized into smaller elements, and numerical solvers are employed in each element to obstain a solution.
A few examples out of these wide range of numerical methods used to solve wave equations are finite differences, finite volumes and finite elements. These numerical methods have played a crucial role in advancing the field of earthquake simulation and are continuously being improved to produce more accurate and efficient simulations. \\

SeisSol\footnote{\href{https://seissol.org}{https://seissol.org}} is a scientific software that enables the simulation of seismic waves and earthquake phenomena. 
However, the modelling of complex 3D geometries can be computationally intensive. To address this challenge, SeisSol implements an efficient way to optimally utilize the computing resources. 
The software allows for the study of various scenarios, including adjusting material properties to exhibit elastic, viscoelastic and viscoplastic behaviour.
SeisSol utilizes a combination of \acf{DG} method and \acf{ADER} time-integration approach to solve the problems.
This combination is called \ac{ADER}-\ac{DG} approach ~\parencite{dumbser}. ~\parencite{dumbser1} extended this approach to solve the elastic wave equations in three dimensions. The \ac{ADER}-\ac{DG} approach provides arbitrary high-order accuracy in space and time, with the solution in each element approximated by polynomials.
The degrees of freedom are expressed through the coefficients of the polynomial, which are advanced in time during simulation, allowing for discontinuities across element boundaries ~\parencite{martin}. \\

In this thesis, we extend SeisSol with the implementation of Wave Reversal of Seismic waves using the \acf{ITM}. 
Time-reversal techniques have been researched in the field of acoustic media, primarily in order to focus the wave below
the diffraction limit~\parencite{67072}. This methodology may be applied, for instance, to improve the detection of gallstones within the human body~\parencite{10.1063/1.881692}. 
In the context of seismic waves, time-reversal methods may be employed to investigate the source and origin of the seismic waves.
The work by~\parencite{Fink2017} describes two techniques for wave reversal. The first approach is the time-reversal mirror approach, which involves recording the wavefield at some 
receiver locations and reversing the wavefield. This approach has been used experimentally to reverse wavefields in elastic solids~\parencite{PhysRevLett.100.064301}.~\parencite{Weinert2016} describes a Loschmidt's demon as something which reverses the momentum/velocity of each particle instantaneously such that the state of the system is reversed 
from a fibrillated state to a less fibrillated, smoother state. The second approach for wave-reversal is called the \acf{ITM} approach, which involves
mimicking Loschmidt's Demons for waves ~\parencite{Fink2017}. 
While both techniques generate time-reversed waves, we specifically focus on the \ac{ITM} approach. 
This method involves manipulating waves on time boundaries by sudden modifications of the wave properties like wave speed or impedance in the whole medium, 
effectively mimicking the Loschmidt point of view. 
This approach is very efficient for radiating time-reversed waves without the use of any antennas or receivers. 
The first approach, which utilizes "time-reversal mirrors" with wave manipulation along a spatial boundary sampled by a finite number of antennas, has numerous applications in telecommunication, imaging, therapy, and defense ~\parencite{Fink2017}. 
~\parencite{Bacot2016} have experimentally demonstrated the instantaneous time mirrors on surface waves on water by manipulating the wave speed for a period of time using a vertical jolt.
This is shown to refocus the waves at the source position before diverging again.\\

To achieve \ac{ITM}s in a simulation environment, we modify the impedances of seismic waves for a short period of time $t_{ITM}$ to $t_{ITM} + \tau$ by manipulating 
the material properties $\lambda, \mu, \rho$. As the impedances of seismic waves depend on these three properties, this manipulation is 
sufficient to achieve wave reversal using \ac{ITM}s.

It is important to note that the time-reversal presented here, in its simplest form, may not be directly useful in experimental 
investigations, as we cannot directly modify the material properties. However, it does provide an approach to approximately 
locate the source of an earthquake when successfully reversed.\\

The thesis is organized as follows: Chapter \ref{chapter:seismicwaves} focuses on the fundamental theory of elastic wave propagation and the presentation
of the numerical \ac{ADER}-\ac{DG} scheme as used in SeisSol. The equations of motion in elastic media is derived, and the wave equation is presented
in the velocity-stress formulation. Chapter \ref{chapter:TimeReversal} outlines the theory of time-reversal, briefly discussing the wave reversal approach before
delving into a more detailed explanation of the \ac{ITM}s concept. Section \ref{section:Implementation} describes the method and implementation details of \ac{ITM}s
in SeisSol. Chapter \ref{chapter:Results} discusses the different studies and results obtained in the process for different scenarios. Finally, we discuss the conclusions
and possible future research directions in Chapter \ref{chapter:Conclusion}. 

%See~\autoref{tab:sample}, \autoref{fig:sample-drawing}, \autoref{fig:sample-plot}, \autoref{fig:sample-listing}.
% 
% \begin{table}[htpb]
%   \caption[Example table]{An example for a simple table.}\label{tab:sample}
%   \centering
%   \begin{tabular}{l l l l}
%     \toprule
%       A & B & C & D \\
%     \midrule
%       1 & 2 & 1 & 2 \\
%       2 & 3 & 2 & 3 \\
%     \bottomrule
%   \end{tabular}
% \end{table}
% 
% \begin{figure}[htpb]
%   \centering
%   % This should probably go into a file in figures/
%   \begin{tikzpicture}[node distance=3cm]
%     \node (R0) {$R_1$};
%     \node (R1) [right of=R0] {$R_2$};
%     \node (R2) [below of=R1] {$R_4$};
%     \node (R3) [below of=R0] {$R_3$};
%     \node (R4) [right of=R1] {$R_5$};
% 
%     \path[every node]
%       (R0) edge (R1)
%       (R0) edge (R3)
%       (R3) edge (R2)
%       (R2) edge (R1)
%       (R1) edge (R4);
%   \end{tikzpicture}
%   \caption[Example drawing]{An example for a simple drawing.}\label{fig:sample-drawing}
% \end{figure}
% 
% \begin{figure}[htpb]
%   \centering
% 
%   \pgfplotstableset{col sep=&, row sep=\\}
%   % This should probably go into a file in data/
%   \pgfplotstableread{
%     a & b    \\
%     1 & 1000 \\
%     2 & 1500 \\
%     3 & 1600 \\
%   }\exampleA
%   \pgfplotstableread{
%     a & b    \\
%     1 & 1200 \\
%     2 & 800 \\
%     3 & 1400 \\
%   }\exampleB
%   % This should probably go into a file in figures/
%   \begin{tikzpicture}
%     \begin{axis}[
%         ymin=0,
%         legend style={legend pos=south east},
%         grid,
%         thick,
%         ylabel=Y,
%         xlabel=X
%       ]
%       \addplot table[x=a, y=b]{\exampleA};
%       \addlegendentry{Example A};
%       \addplot table[x=a, y=b]{\exampleB};
%       \addlegendentry{Example B};
%     \end{axis}
%   \end{tikzpicture}
%   \caption[Example plot]{An example for a simple plot.}\label{fig:sample-plot}
% \end{figure}
% 
% \begin{figure}[htpb]
%   \centering
%   \begin{tabular}{c}
%   \begin{lstlisting}[language=SQL]
%     SELECT * FROM tbl WHERE tbl.str = "str"
%   \end{lstlisting}
%   \end{tabular}
%   \caption[Example listing]{An example for a source code listing.}\label{fig:sample-listing}
% \end{figure}
% 