% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.

\chapter{1D Acoustic Equation \ac{ITM}}\label{chapter:ITMAcoustic}
In this chapter, we calculate the analytical solution for a 1D Acoustic wave when an \ac{ITM} is applied from $t_{ITM}^-$ to $t_{ITM}^+$.
For this, we consider a 1D Acoustic Equation [~\parencite{leveque_2002}, Sec. 2.8] linearized about the motionless state:
\begin{align}
    \begin{split}
        p_t + K_0u_x = 0, \\
        u_t + \frac{1}{\rho_0}p_x = 0 .\\
    \end{split}
\end{align}

where $p$ is the pressure perturbation, $K_0$ is the bulk modulus of the material and $u$ is the velocity purturbation. 
This can be written in consolidated matrix vector form as done for first order hyperbolic systems of equations as follows:

\begin{align}
    \begin{split}
    \begin{bmatrix}
        p \\
        u
    \end{bmatrix}_t + 
    \begin{bmatrix}
        0 & K_0 \\
        \frac{1}{\rho_0} & 0 \\
    \end{bmatrix}
    \begin{bmatrix}
        p \\
        u
    \end{bmatrix}_x = 0
    \end{split}
    \label{eq:acoustic}
\end{align}.

The flux matrix for equation \ref{eq:acoustic} has two eigen values:

\begin{align}
    \begin{split}
        \lambda_1 = -c_0; \lambda_2 = c_0,
    \end{split}
\end{align}

where 
\begin{align}
    \begin{split}
        c_0 = \sqrt{\frac{K_0}{\rho_0}}
    \end{split}
\end{align}

which is the speed of sound in gas. And waves can propagate in either direction with this speed. The eigen vectors for this coefficient matrix come out to be:

\begin{align}
    \begin{split}
        r^1 = \begin{bmatrix}
            \rho_0 c_0 \\
            1
        \end{bmatrix}, 
        r^2 = \begin{bmatrix}
            \rho_0 c_0 \\
            1
        \end{bmatrix}.
    \end{split}
\end{align}

The impedance is defined as:
\begin{equation}
    Z_0 = \rho_0 c_0.
\end{equation}.

Using these parameters, ~\parencite{leveque_2002} Sec. 2.8 calculates the solution of the acoustic equation depending on the initial conditions $p^0(x), u^0(x)$ as:
\begin{align}
    \begin{split}
        p(x,t) = \frac{1}{2}\left[p^0\left(x + c_0t\right) + p^0\left(x - c_0t\right)\right] - \frac{Z_0}{2}\left[u^0\left(x+c_0t\right) - u^0\left(x-c_0t\right)\right], \\
        u(x,t) = -\frac{1}{2Z_0}\left[p^0\left(x+c_0t\right) - p^0\left(x-c_0t\right)\right] + \frac{1}{2}\left[u^0\left(x+c_0t\right) + u^0\left(x-c_0t\right)\right] .
    \end{split}
    \label{eq:solutionacoustic}
\end{align}

We use this solution to find the analytical solution of a 1D acoustic wave after the \ac{ITM} is applied in two different ways. First we ensure we use initial conditions which give rise to only one forward wave. We simply use a sinusoidal wave multiplied to only one eigen vector $r^2$.
This defines our initial conditions as:
\begin{align}
    \begin{split}
        p^0\left(x\right) = \rho_0c_0cos\left(k*x\right), \\
        u^0\left(x\right) = cos\left(k*x\right) .
    \end{split}
    \label{eq:initialconditions}
\end{align}

where $k$ is the wave number of the wave. The energy of the system is defined as:

\begin{align}
    \begin{split}
     E = \int_{-\frac{\pi}{k}}^{\frac{\pi}{k}} \left(\frac{1}{2K}p^2 + \frac{1}{2}\rho u^2\right) dx , \\
     \implies E = \frac{\pi\rho_0}{k} .
    \end{split}
\end{align}

\section{Impedance Changing \ac{ITM}}
Using the conditions given in equation \ref{eq:initialconditions}, we calculate the solution at $t_ITM^-$ and use them as initial conditions for the new equation where the wave velocities are scaled.
In this case, we scale the velocities such that the impedance also changes(keeping the density constant). Using \ref{eq:solutionacoustic}, we get the initial conditions for the next phase of the simulation$\left(t_{ITM}^- \leq t \leq t_{ITM}^+ \right)$ at $t_{ITM}^-$ as:

\begin{align}
    \begin{split}
        p^0_{t_{ITM}^-}\left(x\right) = c_0 \rho_0 cos\left(kx - c_0kt_{ITM}^-\right), \\
        u^0_{t_{ITM}^-}\left(x\right) = cos\left(kx - c_0kt_{ITM}^-\right) .
    \end{split}
\end{align}

We now use these as the initial conditions and equation \ref{eq:solutionacoustic} with the updated wave velocity and constant density to get the evolved solution for $t_{ITM}^- \leq t \leq t_{ITM}^+ $. ~\parencite{sagemath} was used to perform the simplifications for the arithmetic. We obtain the solution for this period as:
